INGRID: Creating Languages in MPS from ANTLR Grammars

### 2016-10-13 chapters/background.tex
writing about related projects

\paragraph{BaseLanguage.}
Another important feature of MPS that we need to describe in more detail is \emph{BaseLanguage}.
It is a clone of Java implemented using the MPS constructs.
BaseLanguage was developed in the early days of MPS in order to implement MPS itself and also to support the basic set of language-definition DSLs,
Although BaseLanguage is syntactically almost identical to Java, it is edited in a projectional editor, just like all the languages in MPS.
The language-definition DSLs, used to define custom languages by their authors, are generated into the BaseLanguage.
Similarly, all the custom DSLs that are meant to be generated into Java choose BaseLanguage as their generation target.
The conversion to textual Java is handled by BaseLanguage without any further manual effort, since BaseLanguage has a TextGen aspect defined, which translates BaseLanguage code into textual Java sources.

@@

@@

% TODO mozna zmenit nazev teto sekce
\subsection{Similar Projects}

We are aware of several projects that provide certain support for GPLs in the context of the MPS platform.
We will analyze existing related projects that are trying to accomplish similar goals, and then we will weigh advantages and disadvantages of these approaches.

Currently, there exists an almost full port of the Java language called BaseLanguage~\cite{ref:BaseLanguage} extended with some MPS specific features.
It was imported manually by JetBrains and it is still undergoing changes as Java itself is evolving.

Within the mbeddr project~\cite{ref:mbeddr}, the C language is also manually tailored for MPS.

Now/here we describe three similar projects (with similar goals) that we are aware of.
We focus on (consider) these three projects: PE4MPS, ANTLR{\_}MPS, and mps-metabnf.

>>>>

\paragraph{BaseLanguage.}
Another important feature of MPS that we need to describe in more detail is \emph{BaseLanguage}~\cite{ref:BaseLanguage}.
It is a clone of Java implemented using the MPS constructs.
BaseLanguage was developed in the early days of MPS in order to implement MPS itself and also to support the basic set of language-definition DSLs,
Although BaseLanguage is syntactically almost identical to Java, it is edited in a projectional editor, just like all the languages in MPS.
The language-definition DSLs, used to define custom languages by their authors, are generated into the BaseLanguage.
Similarly, all the custom DSLs that are meant to be generated into Java choose BaseLanguage as their generation target.
The conversion to textual Java is handled by BaseLanguage without any further manual effort, since BaseLanguage has a TextGen aspect defined, which translates BaseLanguage code into textual Java sources.

@@

\subsection{Related Projects}

The BaseLanguage~\cite{ref:BaseLanguage} is an almost full port of the Java language, extended with MPS-specific features.
It was created manually by JetBrains, and it is still undergoing changes as Java itself is evolving.

The C language has also been manually tailored for MPS within the mbeddr project~\cite{ref:MBEDDR}.

Besides that, we are aware of three other projects that provide certain support for GPLs in the context of MPS --- namely PE4MPS, ANTLR{\_}MPS, and mps-metabnf.
We describe their main features and limitations in this section.

@@

### 2016-10-13 paper.tex
finished introduction
tweak code listing
structure
editor and textgen
writing about related projects

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{url}
\usepackage{xspace}
\usepackage{listings}
\usepackage[noline]{algorithm2e}
\usepackage{comment}
\usepackage{color}
\usepackage{times}
\usepackage{multirow}
\usepackage{balance}
\usepackage{adjustbox}
\usepackage{alltt}
\usepackage{textcomp}
\usepackage[usenames]{xcolor}

@@

% colors for the ANTLR environment
\definecolor{antlrbgcolor}{rgb}{1,1,1}
\definecolor{antlrparserrulecolor}{RGB}{11,83,148}
\definecolor{antlrlexerrulecolor}{RGB}{191,144,0}
\definecolor{antlrliteralcolor}{RGB}{133,32,12}
\definecolor{antlrregexcolor}{RGB}{56,118,29}

% colors for the MPS environment
\definecolor{mpsbgcolor}{rgb}{1,1,1}
\definecolor{mpsconceptcolor}{RGB}{42,140,0}
\definecolor{mpsinterfacecolor}{RGB}{170,0,229}

% coloring commands
\newcommand{\antlrap}{\textquotesingle}
\newcommand{\antlrparserrule}[1]{\textcolor{antlrparserrulecolor}{#1}}
\newcommand{\antlrlexerrule}[1]{\textcolor{antlrlexerrulecolor}{#1}}
\newcommand{\antlrliteralnoap}[1]{\textcolor{antlrliteralcolor}{#1}}
\newcommand{\antlrliteral}[1]{\antlrap\antlrliteralnoap{#1}\antlrap}
\newcommand{\antlrregex}[1]{\textcolor{antlrregexcolor}{#1}}
\newcommand{\mpsconcept}[1]{\textcolor{mpsconceptcolor}{#1}}
\newcommand{\mpsinterface}[1]{\textcolor{mpsinterfacecolor}{#1}}

% environment for the ANTLR code (grammars)
\newenvironment{antlr}
{
	\vspace{3mm}
	\par
	\noindent
	\adjustbox{margin=3mm,bgcolor=antlrbgcolor,frame,minipage=485pt,inner=\linewidth}
	\bgroup
	\alltt
}
{
	\endalltt
	\egroup
	\\
	\vspace{2mm}
}

% environment for the MPS code
\newenvironment{mps}
{
	\vspace{3mm}
	\par
	\noindent
	\adjustbox{margin=3mm,bgcolor=mpsbgcolor,frame,minipage=394pt,inner=\linewidth}
	\bgroup
	\alltt
}
{
	\endalltt
	\egroup
	\\
	\vspace{2mm}
}

@@

\title{\todo{INGRID: Creating Languages in MPS from ANTLR Grammars}}

@@

@@

\bibitem{ref:mbeddr} mbeddr, \url{mbeddr.com/}

\bibitem{ref:BaseLanguage} BaseLanguage, \url{confluence.jetbrains.com/display/MPSD34/Base+Language}

@@

>>>>

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{url}
\usepackage{xspace}
\usepackage{listings}
\usepackage[noline]{algorithm2e}
\usepackage{comment}
\usepackage{color}
\usepackage{times}
\usepackage{multirow}
\usepackage{balance}
\usepackage{framed}
\usepackage{alltt}
\usepackage{textcomp}
\usepackage[dvipsnames]{xcolor}

@@

% coloring and font style commands
  % ANTLR
\newcommand{\antlrap}{\textquotesingle}
\newcommand{\antlrparserrule}[1]{\textcolor{Blue}{#1}}
\newcommand{\antlrlexerrule}[1]{\textcolor{RedOrange}{#1}}
\newcommand{\antlrliteralnoap}[1]{\textcolor{Maroon}{#1}}
\newcommand{\antlrliteral}[1]{\antlrap\antlrliteralnoap{#1}\antlrap}
\newcommand{\antlrregex}[1]{\textcolor{PineGreen}{#1}}
  % MPS in general
\newcommand{\mpsconcept}[1]{\textcolor{OliveGreen}{#1}}
\newcommand{\mpsinterface}[1]{\textcolor{Plum}{#1}}
  % MPS structure
\newcommand{\mpsstkeyword}[1]{\textcolor{Blue}{\textbf{#1}}}
\newcommand{\mpsstalias}[1]{\textcolor{Green}{\textbf{#1}}}
\newcommand{\mpsstproperty}[1]{\textcolor{Fuchsia}{\textbf{#1}}}
\newcommand{\mpsstplaceholder}[1]{\textcolor{Gray}{\textbf{#1}}}
\newcommand{\mpsstcardinality}[1]{\textcolor{RoyalBlue}{#1}}
  % MPS editor
\newcommand{\mpsedkeyword}[1]{\textcolor{Black}{\textbf{#1}}}
\newcommand{\mpsedtarget}[1]{\textcolor{Fuchsia}{#1}}
\newcommand{\mpsedannotation}[1]{\textcolor{Gray}{\textbf{#1}}}
  % MPS textgen
\newcommand{\mpstgkeyword}[1]{\textcolor{Blue}{\textbf{#1}}}
\newcommand{\mpstgtarget}[1]{\textcolor{Fuchsia}{#1}}
\newcommand{\mpstgparam}[1]{\textcolor{Violet}{#1}}
\newcommand{\mpstgaction}[1]{\textcolor{Fuchsia}{\textbf{#1}}}
\newcommand{\mpstgliteral}[1]{\textcolor{Green}{\textbf{#1}}}
\newcommand{\mpstgnodeprop}[1]{\textcolor{darkgray}{#1}}

@@

\title{INGRID: Creating Languages in MPS from ANTLR Grammars}

@@

\bibitem{ref:BaseLanguage} MPS BaseLanguage, \url{https://confluence.jetbrains.com/display/MPSD34/Base+Language}

@@

@@

\bibitem{ref:mpsmetabnf} mps-metabnf \url{https://github.com/DSLFoundry/mps-metabnf}
